---
title: "Trials"
author: "Rachel Koerner"
date: "4/28/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Project #2 USGS DOC Data

##--------------------------------------
## Install and load any needed libraries
##--------------------------------------

library(fields)
library(lattice)


##--------------------------------------
## Load the data
##--------------------------------------

load("DOC_baseFlow_weightedAvg_and_predictors.RData")
ls()

#This is a really long name for a dataset, so this command just renames it.

data <- DOC_baseFlow_weightedAvg_and_predictors


#pairs(meanDOC ~., data = data)

```




```{r}
#backwards model
n = nrow(data)
full_model <- lm(meanDOC ~., data = data)
bic <- step(full_model, direction = "backward", k = log(n), trace = 0)
```

The final coefficients are:
  * fire_normRdNBR_5yDecay
  * fire_normRdNBR_10yDecay
  * fire_normRdNBR_15yDecay
  * landCover_developedFraction
  * landCover_forestFraction
  * minTemp_baseFlow
  * maxTemp_baseFlow
  * precip_baseFlow
  * precip_nonBaseFlow
  * SWE_baseFlow
  * soil_Bfraction
  * elevation_mean
  * aspect_predominant
  * wasteWaterPointSources_count
  
```{r}
mic <- step(full_model, direction = "backward", k = n-1, trace = 0)
summary(mic)

```

The final coefficient to be removed is the elevation_mean.

```{r}
plot(data$meanDOC, data$elevation_mean)
model <- lm(elevation_mean ~ meanDOC, data = data)
abline(model, col = "red")
```





```{r}
#forward selection
small_model <- lm(meanDOC ~1, data = data)
fwd <- step(small_model, direction = "forward", scope = formula(full_model), k = log(n), trace = 0)
```

The resulting coefficients are:
  * elevation_mean
  * wasteWaterPointSources_count
  * maxTemp_baseFlow
  * aspect_predominant
  * soil_Bfraction
  * landCover_developedFraction
  * precip_baseFlow
  * landCover_forestFraction
  * fire_normRdNBR_5yDecay
  * minTemp_yearAvg
  * precip_nonBaseFlow




```{r}

#lasso fit
suppressMessages(library(glmnet))
leave_out <- which((colnames(data) == "meanDOC"))
irrelevant <- matrix(rnorm(nrow(data) * 5), nrow(data), 5)
colnames(irrelevant) <- paste("i", 1:5, sep = "")
x <- cbind(as.matrix(data[, -leave_out]), irrelevant)
y <- data$meanDOC

lasso_fit <- cv.glmnet(x, y)
coef(lasso_fit)

```

Coefficients remaining are:
  * maxTemp_baseFlow
  * maxTemp_yearAvg**
  * soil_Afraction**
  * elevation_mean
  * aspect_predominant
  * wasteWaterPointSources_count
  
**unique to lasso fit model


```{r}
plot(lasso_fit)
```




```{r}
plot(data$meanDOC, data$elevation_mean)
model <- lm(elevation_mean ~ meanDOC, data = data)
abline(model, col = "red")

green <- data[data$MPB_greenFraction > 0, ]
red <- data[data$MPB_redFraction > 0, ]
gray <- data[data$MPB_grayFraction > 0, ]


points(red$meanDOC, red$elevation_mean, col = "red")
points(gray$meanDOC, gray$elevation_mean, col = "gray")
points(green$meanDOC, green$elevation_mean, col = "darkgreen")


```


```{r}
plot(data$meanDOC, data$elevation_mean)
model <- lm(elevation_mean ~ meanDOC, data = data)
abline(model, col = "red")

green <- data[data$landCover_developedFraction > 0, ]
red <- data[data$landCover_agriculturalFraction > 0, ]
gray <- data[data$landCover_forestFraction > 0, ]

points(gray$meanDOC, gray$elevation_mean, col = "gray")
points(green$meanDOC, green$elevation_mean, col = "darkgreen")
points(red$meanDOC, red$elevation_mean, col = "red")


```